### 2012-11-06 paper_sean/hyperneat.tex
added changes we talked about today..near final?

In this study, the ANN inputs, outputs, activation functions, and the size of the hidden layer are the same as in Yosinski et al.~\cite{yos:clune}. 
The ANN had a fixed topology of three \begin{math}3\times4\end{math} Cartesian grids of nodes for the input, hidden, and output layers. 
The inputs to the substrate were the angles requested in the prevous time step for each of the 9 joints of the robot and a sine and cosine wave to facilitate periodic motion. 
The outputs of the substrate at each time step were nine numbers (for each joint) in the range [-1, 1] which were scaled to the range [0, 1023], the allowable ranges for the servos. 
As in Yosinski et al.~\cite{yos:clune}, we generated pseudo-positions at 160Hz and then downsampled over consecutive blocks of four time steps to obtain the actual commanded positions at 40Hz; this reduced the number of gaits which commanded switches from extreme negative to extreme positive numbers at 40Hz, which overly taxed the servos.

>>>>

In this study, the ANN inputs, outputs, activation functions, and the size of the hidden layer are the same as in Yosinski et al.~\cite{yos:clune}. 
The ANN had a fixed topology of three \begin{math}3\times4\end{math} Cartesian grids of nodes for the input, hidden, and output layers. 
The inputs to the substrate were the angles requested in the previous time step for each of the 9 joints of the robot and a sine and cosine wave to facilitate periodic motion. 
The outputs of the substrate at each time step were nine numbers (for each joint) in the range [-1, 1] which were scaled to the allowable ranges for the servos. 
As in Yosinski et al.~\cite{yos:clune}, we generated pseudo-positions at 160Hz and then downsampled over consecutive blocks of four time steps to obtain the actual commanded positions at 40Hz; this reduced the number of gaits which commanded switches from extreme negative to extreme positive numbers at 40Hz, which overly taxed the servos.

### 2012-11-06 paper_sean/results.tex
added changes we talked about today..near final?

The previous fastest gait on any copy of a QuadraTot was from Glette
et al., and traveled 17.8 cm/s on the ROBIN QuadraTot. The
fastest gait produced by HyperNEAT with a simulator in the experiments for this paper traveled 14.5
cm/s on the CCML QuadraTot. It was unclear whether this difference in performance was due to
the differences in the gaits themselves or dissimilarities between
hardware. To control for this possibility, we ran the fastest gait
from Glette et al. 10 times on CCML and measured an average speed of
only 12.95 cm/s (vs. 17.8 cm/s measured on ROBIN) and a maximum of
13.8 cm/s (\tabref{results}). It thus appears that the CCML version of the robot is slower. Moreover, the HyperNEAT gait (14.5 cm/s) is faster than the best gait of Glette et al. on the CCML QuadraTot. Unfortunately, it was not possible to test the best HyperNEAT gait on the ROBIN QuadraTot. Because HyperNEAT outperformed the simple encoding from Glette et al. on the same robotic hardware, we tentatively
conclude that HyperNEAT produces faster gaits for the QuadraTot robot. This conclusion is supported by the fact that HyperNEAT also outperformed the encoding from Glette et al. in simulation.

\begin{table}
\begin{center}
\begin{tabular}{|l|r|r|r|r|}
\hline
                                         &              & Simulated  & Real Vel. & Real Vel.  \\
                                         & Evaluations  & Velocity &    (CCML)     & (ROBIN) \\
\hline
Parameterized gaits + optimization \cite{yos:clune}   &153    & --    & 5.8 & --\\
\hline
HyperNEAT in hardware \cite{yos:clune}                 & 180         & --         &   9.7  & --   \\
\hline
RL PoWER Spline \cite{haocheng}                         & 300         & --         &   11.1 & --\\
\hline
GA + simulator \cite{glette}             & 60000       & *16.7       &   13.8   & \textbf{17.8}  \\
\hline
HyperNEAT + simulator [this paper]                     & 40000       & \textbf{25.4}       &   \textbf{14.5} & --\\
\hline
\end{tabular}
\vspace{.35cm}
\caption{Velocities of evolved gaits in simulation and on two different copies of the QuadraTot robot. Subject to availability, data are reported from the experiments for this paper and three previous studies. Reported are the total number of evaluations per run, the average of the fastest gaits produced in each run in simulation, and the fastest gait produced on the CCML and ROBIN copies of the QuadraTot robot (see text). *Instead of using the single fitness value reported in~\cite{glette}, we ran 19 additional runs and used the average fitness of those 20 runs. Velocities are in cm/s, and bold indicates the best performance.}  
\tablabel{results}
\end{center}
\end{table}

>>>>

The previous fastest gait on any copy of a QuadraTot was from Glette
et al., and traveled 17.8 cm/s on the ROBIN QuadraTot. The
fastest gait produced by HyperNEAT with a simulator in the experiments for this paper traveled 14.5
cm/s on the CCML QuadraTot. It was unclear whether this difference in performance was due to
the differences in the gaits themselves or dissimilarities between
hardware. To control for this possibility, we ran the fastest gait
from Glette et al. 10 times on CCML and measured an average speed of
only 12.95 cm/s $\pm0.93$ SD (vs. 17.8 cm/s measured on ROBIN) and a maximum of
13.8 cm/s (\tabref{results}). It thus appears that the CCML version of the robot is slower. Moreover, the HyperNEAT gait (14.5 cm/s) is faster than the best gait of Glette et al. on the CCML QuadraTot. Unfortunately, it was not possible to test the best HyperNEAT gait on the ROBIN QuadraTot. Because HyperNEAT outperformed the simple encoding from Glette et al. on the same robotic hardware, we tentatively
conclude that HyperNEAT produces faster gaits for the QuadraTot robot. This conclusion is supported by the fact that HyperNEAT also outperformed the encoding from Glette et al. in simulation.

\begin{table}
\begin{center}
\begin{tabular}{|l|r|r|r|r|}
\hline
                                         &              & Simulated  & Real Vel. & Real Vel.  \\
                                         & Evaluations  & Velocity &    (CCML)     & (ROBIN) \\
\hline
Parameterized gaits + optimization \cite{yos:clune}   &153    & --    & 5.8 & --\\
\hline
HyperNEAT in hardware \cite{yos:clune}                 & 180         & --         &   9.7  & --   \\
\hline
RL PoWER Spline \cite{haocheng}                         & 300         & --         &   11.1 & --\\
\hline
GA + simulator \cite{glette}             & 60000       & *16.7       &   13.8   & \textbf{17.8}  \\
\hline
HyperNEAT + simulator [this paper]                     & 40000       & \textbf{25.4}       &   \textbf{14.5} & --\\
\hline
\end{tabular}
\vspace{.35cm}
\caption{Velocities of evolved gaits in simulation and on two different copies of the QuadraTot robot. Subject to availability, data are reported from the experiments for this paper and three previous studies. Reported are the total number of evaluations per run, the average of the fastest gaits produced in each run in simulation, and the single fastest gait produced on the CCML and ROBIN copies of the QuadraTot robot (see text for their differences). *Instead of using the single fitness value reported in~\cite{glette}, we ran 19 additional runs and used the average fitness of those 20 runs. Velocities are in cm/s, and bold indicates the best performance. }  
\tablabel{results}
\end{center}
\end{table}
